---
title: "USDA NASS Soybeans data"
subtitle: "Geospatial summaries"
author: "All of us"
format:
  html:
    code-fold: true
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
    #cerulean, journal, quartz, sketchy, slate, morph, superhero, vapor
knitr: 
  opts_chunk: 
    message: false
    warning: false
---

# Introduction  
This report was developed to explore, through **maps**, USDA NASS data for **soybeans production at the county level for all years available**.  

The raw data was downloaded from [USDA NASS QuickStats portal](https://quickstats.nass.usda.gov).  

# Learning objectives  
The learning objectives for this exercise are:  
- Import wrangled USDA NASS data  
- Connect tabular data with spatial data based on county information  
- Create a pub ready static map  
- Create an interactive map  
- Explore a rendered quarto script with interactivity  

# Setup  
## Loading packages  
```{r }
#install.packages("remotes")
#remotes::install_github("ropensci/USAboundaries")
#remotes::install_github("ropensci/USAboundariesData")
#install.packages("mapview")
#install.packages("viridis")

library(tidyverse)
library(janitor)
library(ggthemes)
library(sf)
library(USAboundaries)
library(mapview)
library(viridis)

```

## Importing data  
```{r soy}
soy_w 

soy_w
```

## Creating geospatial data 
```{r}
ga_co 

# Checking class
class(ga_co)

# Exploring the object
ga_co 

# Quick plot

```

Our goal now is to merge both the NASS data set `soy_w` with the GA counties geospatial dataset `ga_co`.  

# Wrangling  
## NASS data  
```{r soy_w_all}
soy_w_all <- soy_w %>%
  # Removing other counties
  filter(county != "OTHER COUNTIES" &
           county != "OTHER (COMBINED) COUNTIES") %>%
  # Matching county capitalization to ga_co
  mutate(county = str_to_title(county)) 

soy_w_all 
```

## County geospatial data  
```{r ga_co_w}
ga_co_w 

ga_co_w    
```

## Joining NASS and county  
```{r soy_sp}
soy_sp  

soy_sp
```

- Check number of rows of product, compare to `soy_w_all`  

# Maps - static  
## Plot 2023 production by county as a map  
```{r}
soy_sp 
```

## Make map above publication ready  
```{r}
soy_sp %>%
  filter(year == 2024) %>%
  ggplot()+
  geom_sf(data = ga_co_w,
          size = .2)+
  geom_sf(data = ~., aes(fill = prod_mibu),
          size = .2
  )
  # Selecting a colorblind-safe palette

  # Changing fill legend label, giving a title

  # Adding county names

  # Changing to a cleaner theme


```
# Maps - interactive  
## Interactive map of most recent year
```{r}
soy_sp_22 <- soy_sp %>%
          filter(year == 2024)

```

# Summary  
In this exercise we  
- Learned how to connect tabular and spatial data  
- Create pub ready static map, and an interactive map  

