---
title: "Digitizing field boundary"
format: html
---

# Introduction  
When working with geospatial data, having a boundary of a field is one of the first steps.  

When working with on-farm data, it is common for growers to already have a boundary file, most likely in a shapefile (.shp) format.  

In case they don't, the easiest way to share that with you is by providing lon/lat coordinates, which are easily retrieved on Google maps or any other mobile phone map app.  

If they do share lon/lat coordinates and ask for you to create a boundary, then you would need to know how to do so on your side.  

The act of creating a field boundary is referred to as **digitizing**.  

Digitizing a field boundary can be accomplished in any GIS software including ArcMap, QGIS, Google Earth, and R.  

Let's explore how to digitize in R in this exercise.  

# Situation  
Throughout the semester, we will start working with data from a field in south Georgia.  

Suppose this grower does not have a boundary file, and asks you to create it.  

All they provide you with are the lon/lat coordinates of the field centroid:  

> lon: -84.76151  
> lat: 31.18928
 
Let's learn how to digitize the boundary for this field.  
# Setup  
```{r loading packages}
library(tidyverse)
library(sf)
library(mapview)
library(leaflet.extras) # for digitizing drawing tool bar  
library(mapedit) # for creating the boundary

```

# Wrangling  

We need to transform the coordinates into a spatial point.  

```{r centroid}
centroid 
  # Transforming df into sf

  # Assigning proper CRS


```
Let's visualize it with `mapview()`  

```{r}

```

Ok, we have an idea of where the field is based on its centroid.  

Now, let's go ahead and create the boundary of this field.  

# Digitizing boundary  
First, we need to create a map call that allows for digitizing, as the default behavior of mapview alone does not provide this option.  

```{r}
# Creating map for digitizing
digitizing 

digitizing
```

Now, we initialize a map view session, and we are done with the boundary and hit `Done`, then the digitized boundary will be assigned to an R object.  

Steps:  
1. Run the code in the chunk below  

2. A Viewer window will appear. Pop it up by clicking on the up-arrow symbol  

3. Once the map is up, on the left-hand side there is a drawing menu. Click on the option for **"Draw a polygon"**.  

4. Click on the map to start adding points to your boundary.  

5. Keep adding points to the boundary to draw its shape. You may need to zoom out and back in to adjust your view of the field.  

6. To finalize the digitizing, just click again on the first point to close the shape.  

7. To make sure your digitized boundary is "saved" to an object, click on the "Done" button on the right bottom side of the pane. This will close the viewer window.  


```{r}
boundary 

boundary
```

Inspect the `boundary` object after you are done.  
It has `$drawn`, `$edited`, `$deleted`, `$finished` and `$all` list components.  

We want to retrieve the `$finished` component.  

```{r}
boundary_w  

boundary_w
```
Let's inspect it and make sure we got it correctly.  
```{r}

```
How does it look? If you are satisfied with the result, proceed to exporting it.  

If something is not up to par, then redo the digitizing steps.  

# Exporting boundary  
Now all we need to do is export it to file, so we can use it for all the upcoming analysis.  

Notice that as of now, the boundary is assigned to an object named `boundary_w`, and this object is of type `sf`.  

We need to export it to file and provide it with a sensible file extension.  

A couple of options would be as shapefile (`.shp`), or geojson (`.geojson`).  

Let's do both and talk about their differences.

First, as shapefile
```{r}

```

Go to the `output` folder and check what happened.  

This is how shapefiles are exported.  
The `.shp` file comes along with other accompanying files like `.dbf`, `.prj`, and `.shx`, where each contains a different piece of information needed to use the `.shp` file.  

Shapefiles are the most common geospatial vector types in the industry, but are also cumbersome due to needed all accompanying files to work.  

One alternative is using .geojson:  
```{r}

```

With the `.geojson` extension, only one file was exported, which makes it easier and simpler when sharing with someone else.  

# Summary  
We learned how to digitize a field boundary having only the field centroid coordinates.  

This skill is important as a lot of other geospatial data can only be processed or retrieved through having an accurate field boundary of the field.  

